---
title: "SSTI 090 Report - Results"
author: "Ryan Johnson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(drake)
library(tidyverse)
library(RColorBrewer)
library(phyloseq)

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.height = 5,
	fig.width = 9
)
theme_set(theme_minimal())
plot_color <- colorRampPalette(brewer.pal(8, "Set1"))(12)
```

# Relative Abundance

```{r}
# Read in phylum relative abundance
phylum_relabundance <- psmelt(readd(ps_phylum_relabun))

# Read in genus relative abundance
genus_relabundance <- psmelt(readd(ps_genus_relabun))

# Functions ----------------------------------------------------
# Stacked Bar plot
stacked_bar_plot_phylum <- function(body_site){
  phylum_relabundance %>% 
    filter(Body_Site == body_site) %>% 
    ggplot(aes(x = Sample, y = Abundance, fill = phylum)) +
    geom_bar(stat = "identity") +
    labs(x = "",
         y = "Relative Abundance") +
    facet_wrap(~ Day, scales = "free") +
    scale_fill_manual(values = plot_color, na.value = "black") +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 12),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 12)
  )
}
# Phylum Box Plot
boxplot_phylum <- function(body_site){
  phylum_relabundance %>% 
    filter(Sample_Group %in% c("Control", "SSTI")) %>% 
    filter(Body_Site == body_site) %>% 
    ggplot(aes(x = phylum, y = Abundance, fill = Sample_Group)) +
    geom_boxplot() +
    ylim(0, 100) +
    labs(x = "",
         y = "Relative Abundance") +
    facet_wrap(~ Day, scales = "free") +
    theme(
      axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
      axis.text.y = element_text(size = 12),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 12)
  )
}
  
# Get top N genera
top_n_genera <- function(body_site, N){
  genus_relabundance %>% 
    filter(Body_Site == body_site) %>% 
    group_by(genus) %>% 
    summarise(mean_genus_abun = mean(Abundance), .groups = "drop") %>% 
    arrange(desc(mean_genus_abun)) %>% 
    head(N) %>% 
    pull(genus)
}

# Genus Relabun
stacked_bar_plot_genus <- function(body_site, N){
  genus_relabundance %>% 
    filter(Body_Site == body_site) %>% 
    filter(genus %in% top_n_genera(body_site, N)) %>% 
    ggplot(aes(x = Sample, y = Abundance, fill = genus)) +
    geom_bar(stat = "identity") +
    labs(x = "",
         y = "Relative Abundance") +
    scale_fill_manual(values = plot_color, na.value = "black") +
    facet_wrap(~ Day, scales = "free") +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 12),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 12)
  )
}

# Alpha Diversty box plots
alpha_boxplot <- function(metric = c("Observed", "Shannon", "Chao1",
                                    "Simpson", "InvSimpson")){
  metric_sym <- sym(metric)
  alpha_diversity_stats %>% 
    filter(Sample_Group %in% c("Control", "SSTI")) %>% 
    ggplot(aes(x = factor(Day), y = !!metric_sym, fill = Body_Site)) +
    geom_boxplot() +
    labs(x = "Day",
         y = metric) +
    theme(
      axis.text.x = element_text(size = 12, vjust = 0, hjust = 0.5),
      axis.text.y = element_text(size = 12),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 12)
      )
}

# Alpha Diversity box plots: SSTI vs Control
alpha_boxplot_ssti_ctrl <- function(body_site, metric = c("Observed", "Shannon", "Chao1",
                                    "Simpson", "InvSimpson")){
  metric_sym <- sym(metric)
  alpha_diversity_stats %>% 
    filter(Sample_Group %in% c("Control", "SSTI")) %>% 
    filter(Body_Site == body_site) %>% 
    ggplot(aes(x = Day, y = !!metric_sym, fill = Sample_Group)) +
    geom_boxplot() +
    labs(x = "Day",
         y = metric) +
    theme(
      axis.text.x = element_text(size = 12, vjust = 0, hjust = 0.5),
      axis.text.y = element_text(size = 12),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 12)
      )
}

# Change in Diversity (line graph)
change_div_ssti_ctrl <- function(metric){
  metric_sym <- sym(metric)
  tibble(alpha_diversity_stats) %>% 
    filter(Sample_Group %in% c("Control", "SSTI")) %>%
    filter(!is.na(Day)) %>% 
    select(Study_ID, Sample_Group, !!metric_sym, Day, Body_Site) %>% 
    pivot_wider(names_from = Day, values_from = !!metric_sym) %>% 
    drop_na() %>% 
    mutate(Day_0 = 0,
           Day_14 = `14` - `0`,
           Day_28 = `28` - `14`,
           Day_56 = `56` - `28`,
           Day_90 = `90` - `56`) %>% 
    select(-c(`0`, `14`, `28`, `56`, `90`)) %>% 
    pivot_longer(!Study_ID:Body_Site, names_to = "Day", values_to = "Change") %>% 
    group_by(Body_Site, Sample_Group, Day) %>%
    summarise(mean_change = mean(Change), sd = sd(Change)) %>% 
    ggplot(aes(x = Day, y = mean_change, group = Body_Site)) +
    geom_line(aes(color = Body_Site)) +
    geom_point(aes(color = Body_Site)) +
    facet_wrap(~Sample_Group)
}
```


## Phylum Relative Abundance by Day {.tabset}

### Nares

```{r fig.height=5, fig.width=9}
stacked_bar_plot_phylum("Nares")
```

### Oropharynx

```{r fig.height=5, fig.width=9}
stacked_bar_plot_phylum("Oropharynx")
```

### Inguinal

```{r fig.height=5, fig.width=9}
stacked_bar_plot_phylum("Inguinal")
```

### Perianal

```{r fig.height=5, fig.width=9}
stacked_bar_plot_phylum("Perianal")
```

### Abscess

```{r fig.height=5, fig.width=9}
stacked_bar_plot_phylum("Abscess")
```


## Phylum Relative Abundance by Day: SSTI vs Controls {.tabset}

### Nares

```{r fig.height=5, fig.width=9}
boxplot_phylum("Nares")
```

### Oropharynx

```{r fig.height=5, fig.width=9}
boxplot_phylum("Oropharynx")
```

### Inguinal

```{r fig.height=5, fig.width=9}
boxplot_phylum("Inguinal")
```

### Perianal

```{r fig.height=5, fig.width=9}
boxplot_phylum("Perianal")
```




## Genus Relative Abundance by Day {.tabset}

### Nares

```{r fig.height=5, fig.width=9}
stacked_bar_plot_genus("Nares", 10)
```

### Oropharynx

```{r fig.height=5, fig.width=9}
stacked_bar_plot_genus("Oropharynx", 10)
```

### Inguinal

```{r fig.height=5, fig.width=9}
stacked_bar_plot_genus("Inguinal", 10)
```

### Perianal

```{r fig.height=5, fig.width=9}
stacked_bar_plot_genus("Perianal", 10)
```

### Abscess

```{r fig.height=5, fig.width=9}
stacked_bar_plot_genus("Abscess", 10)
```


# Alpha Diversity

```{r}
alpha_diversity_stats <- readd(ps_alpha) %>% 
  mutate(sample_name = readd(sample_names)) %>% 
  full_join(readd(ssti_metadata), by = "sample_name")
```

## Body site diversity levels over time {.tabset}

### Observed ASVs

```{r fig.height=5, fig.width=9}
alpha_boxplot("Observed")
```

### InvSimpson

```{r fig.height=5, fig.width=9}
alpha_boxplot("InvSimpson")
```

### Shannon

```{r fig.height=5, fig.width=9}
alpha_boxplot("Shannon")
```

### Chao1

```{r fig.height=5, fig.width=9}
alpha_boxplot("Chao1")
```


## Alpha Diversity: SSTI vs Control {.tabset}

### Nares {.tabset .tabset-pills}

#### Observed ASVs

```{r}
alpha_boxplot_ssti_ctrl("Nares", "Observed")
```

#### InvSimpson

```{r}
alpha_boxplot_ssti_ctrl("Nares", "InvSimpson")
```

#### Shannon

```{r}
alpha_boxplot_ssti_ctrl("Nares", "Shannon")
```

#### Chao1

```{r}
alpha_boxplot_ssti_ctrl("Nares", "Chao1")
```

### Inguinal {.tabset .tabset-pills}

#### Observed ASVs

```{r}
alpha_boxplot_ssti_ctrl("Inguinal", "Observed")
```

#### InvSimpson

```{r}
alpha_boxplot_ssti_ctrl("Inguinal", "InvSimpson")
```

#### Shannon

```{r}
alpha_boxplot_ssti_ctrl("Inguinal", "Shannon")
```

#### Chao1

```{r}
alpha_boxplot_ssti_ctrl("Inguinal", "Chao1")
```

### Oropharynx {.tabset .tabset-pills}

#### Observed ASVs

```{r}
alpha_boxplot_ssti_ctrl("Oropharynx", "Observed")
```

#### InvSimpson

```{r}
alpha_boxplot_ssti_ctrl("Oropharynx", "InvSimpson")
```

#### Shannon

```{r}
alpha_boxplot_ssti_ctrl("Oropharynx", "Shannon")
```

#### Chao1

```{r}
alpha_boxplot_ssti_ctrl("Oropharynx", "Chao1")
```

### Perianal {.tabset .tabset-pills}

#### Observed ASVs

```{r}
alpha_boxplot_ssti_ctrl("Perianal", "Observed")
```

#### InvSimpson

```{r}
alpha_boxplot_ssti_ctrl("Perianal", "InvSimpson")
```

#### Shannon

```{r}
alpha_boxplot_ssti_ctrl("Perianal", "Shannon")
```

#### Chao1

```{r}
alpha_boxplot_ssti_ctrl("Perianal", "Chao1")
```

## Mean Change in Diversity levels over time {.tabset}

### Observed ASVs

```{r}
change_div_ssti_ctrl(metric = "Observed")
```

### InvSimpson

```{r}
change_div_ssti_ctrl(metric = "InvSimpson")
```

### Shannon

```{r}
change_div_ssti_ctrl(metric = "Shannon")
```

### Chao1

```{r}
change_div_ssti_ctrl(metric = "Chao1")
```